\documentclass[11pt,twoside,letterpaper,openany]{book}
\setcounter{tocdepth}{4}

\input{packages}
\input{commands}
\usepackage{enumerate}


\begin{document}

\frontmatter

	%% title page
	\begin{titlepage}
	\raggedright\thispagestyle{empty}
	{\Huge Undergraduate Fundamentals \\ of Machine Learning}
	\vspace{\stretch{1}}

	\begin{center}
		{ William J. Deuschle} \\
	\end{center}
	\begin{center}
		{ Adviser: Finale Doshi-Velez}
	\end{center}
	\vspace{\stretch{0.75}}
	\begin{center}
		A thesis presented to Computer Science in partial fulfillment of the honors requirement for the degree of Bachelor of Arts \\
	\end{center}
	\vspace{\stretch{0.75}}

	{\Large Harvard College \\ Cambridge, Massachusetts \\ December, 2018\\}
	\end{titlepage}

	\newpage
	\begin{align*}
		\textbf{Abstract}
	\end{align*}
	Drawing on lectures, course materials, existing textbooks, and other resources, we synthesize and consolidate the content necessary to offer a successful first exposure to machine learning for students with an undergraduate-level background in linear algebra and statistics. The final product is a textbook for Harvard's introductory course in machine learning, CS 181. \newline \newline
	This work is motivated by a lack of resources for individuals with an undergraduate background in the areas necessary to succeed in an introductory course in machine learning. Specifically, existing textbooks are too encyclopedic to be expedient for students seeing the material for the first time, or they assume mathematical and statistical maturity beyond that of undergraduates. Like other professors and students have done for several courses at Harvard, we seek to solve this problem with a properly-scoped textbook following the trajectory of the course.

	\newpage
	\begin{align*}
		\textbf{Acknowledgements}
	\end{align*}
	Much graditude is owed to my thesis adviser, Prof. Finale Doshi-Velez, for patiently guiding me through both the technical and teaching-focused aspects of this thesis. \newline \newline
	I'd also like to thank my family, particularly my parents, for their constant support and encouragement throughout my life, as well as the many friends who aided me while writing this thesis.

	%% table of contents
	\begin{onehalfspacing}
	\tableofcontents
	\end{onehalfspacing}

\mainmatter
    \newchapter{1}{../MLIntroduction/MLIntroduction.tex}{Introduction to Machine Learning}
    \newchapter{2}{../LinearRegression/LinearRegression.tex}{Regression}
    \newchapter{3}{../Classification/Classification.tex}{Classification}
    \newchapter{4}{../NeuralNetworks/NeuralNetworks.tex}{Neural Networks}
    \newchapter{5}{../SupportVectorMachines/SupportVectorMachines.tex}{Support Vector Machines}
    \newchapter{6}{../Clustering/Clustering.tex}{Clustering}
    \newchapter{7}{../DimensionalityReduction/DimensionalityReduction.tex}{Dimensionality Reduction}
    \newchapter{8}{../GraphicalModels/GraphicalModels.tex}{Graphical Models}
    \newchapter{9}{../MixtureModels/MixtureModels.tex}{Mixture Models}
	\newchapter{10}{../HiddenMarkovModels/HiddenMarkovModels.tex}{Hidden Markov Models}
	% \newchapter{11}{../MarkovDecisionProcesses/MarkovDecisionProcesses.tex}{Markov Decision Processes}
	% \newchapter{12}{../ReinforcementLearning/ReinforcementLearning.tex}{Reinforcement Learning}

\end{document}
